\documentclass[a4paper,11pt]{article}

\usepackage{a4wide}
\usepackage{palatino}

\DeclareFixedFont{\titlefont}{OT1}{phv}{b}{n}{16pt}
\DeclareFixedFont{\titlefontsmall}{OT1}{phv}{m}{n}{14pt}
\DeclareFixedFont{\headingfont}{OT1}{phv}{b}{n}{12pt}
\DeclareFixedFont{\subheadingfont}{OT1}{phv}{b}{n}{11pt}
\DeclareFixedFont{\subsubheadingfont}{OT1}{phv}{m}{n}{11pt}

\newcommand{\heading}[1]{\vspace*{8pt}{\headingfont #1}}
\newcommand{\subheading}[1]{\vspace*{8pt}{\subheadingfont #1}}
\newcommand{\subsubheading}[1]{\vspace*{8pt}{\subsubheadingfont #1}}

\setlength{\parskip}{1.5ex}
\setlength{\parindent}{0pt}

\begin{document}

{\titlefontsmall Cambridge University Caving Club}\smallskip\\
{\titlefont Constitution}\medskip\\

\begin{enumerate}
\item This organisation shall be called the ``Cambridge University Caving Club''.

\item The object of the Club shall be to promote active interest in all aspects of
speleology.

\item {\bf Membership}

\begin{enumerate}

\item The membership of the Club shall be open to all past and present members
of the University of Cambridge, students pursuing a course of full-time study
at a college of higher education in Cambridge or at a research institute
approved by the University of Cambridge, employees of the University of
Cambridge or of a college of higher education in Cambridge, and other people
accepted by the committee, regardless of religion, nationality, ethnicity,
gender, sexual orientation, stature, lardiness or lack of personal hygiene. The
words ``in Cambridge'' shall be taken to mean ``within fifteen miles of
Great St. Mary's''. 

\item Honorary members may only be elected by a two-thirds or greater
majority vote at a general meeting. Honorary membership shall be for
life, on the condition that members provide any details required for the
club's smooth running and fulfil any safety requirements.

\item \sloppy Membership is available until the end of the current calendar
year, on the condition that members provide any details required for the
Club's smooth running, fulfil any safety requirements and pay their
membership fees.

\item Temporary membership is open for 17 consecutive days to anyone
participating for the first time in speleological activities with CUCC,
on the condition that they have provided the necessary details the
Club requires for insurance and safety purposes.

\end{enumerate}

\item {\bf Committee}

\begin{enumerate}
\item The officers of the Club shall consist of:
\label{officers}
\begin{enumerate}
\item Senior Treasurer.

\item The following principal officers, who are also voting committee members:
\begin{itemize}
\item President;
\item Secretary;
\item Junior Treasurer;
\item Tackle Master.
\end{itemize}

\item Up to four other voting committee members.
\end{enumerate}

\item The Senior Treasurer shall be a University teacher, or a senior
administrator or Fellow of a College.

\item The majority of the principal officers must be {\em in statu pupillari}.

\item The committee must have a majority of membership {\em in statu pupillari} or
members of the Regent House.

\item The committee shall have the power to assign posts of authority to the
non-principal committee members, but its four principal officers may only be
reassigned by a majority vote at a general meeting. 

\item Management of the Club shall be the responsibility of the committee. 

\item Five voting committee members, or two-thirds of the voting committee
members, whichever is smaller, including at least three principal officers,
shall constitute a quorum at a committee meeting. 

\item No member may serve on the committee for more than one year without
re-election. 

\item Club members may be co-opted onto the committee by its majority vote to
fill casual vacancies, or as non-voting members. 

\end{enumerate}

\item {\bf Finances}

\begin{enumerate}

\item The Junior Treasurer shall manage the Club funds and produce accounts for
presentation at the AGM for approval.  They shall also submit the previous year's
accounts to the Junior Proctor as and when required. 

\item No officers of CUCC shall be liable for any debt or other obligation of
the Club, except where they have personally authorised it in writing. 

\item Officers of CUCC should not authorise any transaction involving financial
liability, unless they are sure it is {\em intra vires} and that sufficient
funds are available. Funds should only be used for purposes consistent with
this constitution.

\item The signatories to the Club bank accounts shall be the Senior Treasurer,
the Junior Treasurer and the President.

\end{enumerate}

\item {\bf General meetings and elections}

\begin{enumerate}

\item The AGM shall be held in the Easter Term or soon afterwards.  Fourteen
days' notice shall be given to members of the date and the agenda.

\item An EGM may be called by the committee, and shall be called upon the
receipt of a petition signed by at least five Club members. Fourteen days'
notice shall be given to members of the date, which shall be in University Full
Term, and agenda.

\item Members and honorary members are entitled to vote at general meetings.
Temporary members are not entitled to vote.

\item Ten voting members, or one half of the voting membership, whichever is
the smaller, shall constitute a quorum at a general meeting. 

\item The officers of the Club shall be elected at the AGM, or a subsequent EGM
should the AGM not be quorate. 

\item Nominations for the officers of the Club may be made to the Secretary
both at the AGM, and during the two weeks preceding the AGM. 

\item Anyone may be nominated for any number of principal officer positions,
but may only be elected to one, unless a total of fewer than four people
have been nominated for the principal officers, in which case the r\^ole of
Tackle Master may be taken on in addition to another r\^ole.

\item The order of election of the principal officers of the committee shall be
the order in which they are listed in clause \ref{officers}, followed by the election of
other voting members. 

\item A secret ballot shall be held if requested by a voting member. 

\item The election shall be by the majority of votes cast. In the event of a
tie, the retiring President shall have the casting vote. 

\item There shall be no postal or proxy voting at general meetings.

\end{enumerate}

\item {\bf Disbarring members from participating in Club activities}
\begin{enumerate}

\item The committee shall have the right to disbar on grounds of safety any
member or non-member from participation in CUCC's speleological activities. 

\end{enumerate}

\item {\bf Constitutional amendments}
\begin{enumerate}

\item The constitution may be amended at a general meeting by a two-thirds
majority vote, providing that the changes are advertised 14 days before the
general meeting. 

\end{enumerate}

\item {\bf Dissolving the Club}
\begin{enumerate}

\item Should the Club be dissolved, its equipment and funds shall be given to
the Societies Syndicate for the benefit of other outdoor clubs.

\end{enumerate}
\end{enumerate}

\hrulefill

\end{document}

